Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-12 paper.qmd
Updated with new quollr version

```{r}
#| label: code-illustration
# Code to draw illustration for notation
## hexagon binning to have regular hexagons
hb_obj_notation <- hex_binning(
  data = tsne_two_curvy_scaled, 
  bin1 = 7, 
  r2 = 1.6,
  q = 0.1)

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = tSNE1, y = tSNE2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos, cex = 1.5)

@@

algo_obj_two_curvy3 <- gen_nldr_vis_algo_obj(
  high_d_data = training_data_two_curvy, 
  nldr_data = tsne_two_curvy, 
  num_x_bins = num_bins_x_two_curvy)

tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$tr_from_to_df
df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$df_bin_centroids
df_bin_two_curvy3 <- algo_obj_two_curvy3$df_bin

hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
counts_df_two_curvy2 <- algo_obj_two_curvy3$hb_obj$std_cts

@@

hex_grid_coloured_two_curvy3 <-  ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy3, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = tSNE1, y = tSNE2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos, cex = 1.5)
```

@@

predict_two_curvy_obj <- gen_scaled_data(
    data = predict_tsne_two_curvy)

@@

```{r}
#| label: read-two-curvy-clust-nldr
# Read a variety of different NLDR representations of two_non_linear_diff_shaped_close_clusters
# and plot them on same aspect ratio
tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")

nldr_two_non_linear_diff_shaped_close_clusters1 <- tsne_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

tsne_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")

nldr_two_non_linear_diff_shaped_close_clusters6 <- tsne_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

umap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")

nldr_two_non_linear_diff_shaped_close_clusters2 <- umap_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#636363') +
  interior_annotation("b", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

phate_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")

nldr_two_non_linear_diff_shaped_close_clusters3 <- phate_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("c") +
  theme(aspect.ratio = 1)

trimap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")

nldr_two_non_linear_diff_shaped_close_clusters4 <- trimap_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("d") +
  theme(aspect.ratio = 1)

pacmap_two_non_linear_diff_shaped_close_clusters <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")

nldr_two_non_linear_diff_shaped_close_clusters5 <- pacmap_two_non_linear_diff_shaped_close_clusters |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1)
```

```{r}
#| label: combine-data-two_non_linear_diff_shaped_close_clusters

error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
error_two_non_linear_diff_shaped_close_clusters_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
error_two_non_linear_diff_shaped_close_clusters_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
error_two_non_linear_diff_shaped_close_clusters_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
error_two_non_linear_diff_shaped_close_clusters_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
error_two_non_linear_diff_shaped_close_clusters_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")

error_two_non_linear_diff_shaped_close_clusters_tsne2 <- error_two_non_linear_diff_shaped_close_clusters_tsne2 |>
  mutate(method = "tSNE2")

error_two_non_linear_diff_shaped_close_clusters <- bind_rows(error_two_curvy_umap, 
                         error_two_non_linear_diff_shaped_close_clusters_tsne,
                         error_two_non_linear_diff_shaped_close_clusters_phate,
                         error_two_non_linear_diff_shaped_close_clusters_trimap,
                         error_two_non_linear_diff_shaped_close_clusters_pacmap,
                         error_two_non_linear_diff_shaped_close_clusters_tsne2)

error_two_non_linear_diff_shaped_close_clusters <- error_two_non_linear_diff_shaped_close_clusters |>
  mutate(a1 = round(a1, 2)) |>
  filter(bin1 >= 5) |>
  group_by(method, a1) |>
  filter(MSE == min(MSE)) |>
  ungroup()
```

```{r}
#| label: error-comp-two_non_linear_diff_shaped_close_clusters

error_plot_two_non_linear_diff_shaped_close_clusters <-
  plot_mse(error_two_non_linear_diff_shaped_close_clusters) +
  scale_x_continuous(breaks =
    sort(unique(error_two_non_linear_diff_shaped_close_clusters$a1))[
      seq(1, length(
        unique(error_two_non_linear_diff_shaped_close_clusters$a1)), 
        by = 5)]) +
  scale_color_manual(
    values=c('#e41a1c','#ff7f00','#4daf4a', 
             "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() + 
  theme(aspect.ratio = 1.5)
```

@@

error_plot_two_non_linear_diff_shaped_close_clusters + wrap_plots(nldr_two_non_linear_diff_shaped_close_clusters1, nldr_two_non_linear_diff_shaped_close_clusters2, nldr_two_non_linear_diff_shaped_close_clusters3, 
                                    nldr_two_non_linear_diff_shaped_close_clusters4, nldr_two_non_linear_diff_shaped_close_clusters5,
                                    nldr_two_non_linear_diff_shaped_close_clusters6, ncol = 2, 
                                    widths = c(50, 50))
```

@@

two_curvy_tsne_scaled_obj <- gen_scaled_data(
  data = tsne_two_non_linear_diff_shaped_close_clusters)

@@

tsne_two_non_linear_diff_shaped_close_clusters_plt <- tsne_two_curvy_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b1", sc_ltr_pos) +
  theme(aspect.ratio = 1)

## hexagon binning to have regular hexagons
hb_obj_tsne_two_curvy2 <- hex_binning(
  data = tsne_two_curvy_scaled, 
  bin1 = 24, 
  r2 = r2_tsne)

a1_2_tsne <- calc_bins_y(
  bin1 = 24, 
  r2 = r2_tsne
)$a1

@@

## To obtain the standardise counts within hexbins
counts_df4 <- hb_obj_tsne_two_curvy2$std_cts
df_bin_centroids4 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df4, 
  counts_df = counts_df4) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 11")

@@

tsne_data_two_curvy2_with_hb_id <- hb_obj_tsne_two_curvy2$data_hb_id
df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, tsne_data_two_curvy2_with_hb_id)
df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)

@@

## Triangulate bin centroids
tr1_object_two_curvy4 <- tri_bin_centroids(
  df_bin_centroids4, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy4 <- gen_edges(
  tri_object = tr1_object_two_curvy4)

@@

names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

```{r}
#| label: best-fit-pacmap-two-curvy

two_curvy_pacmap_scaled_obj <- gen_scaled_data(
  data = pacmap_two_non_linear_diff_shaped_close_clusters)

pacmap_two_curvy_scaled <- two_curvy_pacmap_scaled_obj$scaled_nldr |>
  mutate(ID = row_number())
lim1 <- two_curvy_pacmap_scaled_obj$lim1
lim2 <- two_curvy_pacmap_scaled_obj$lim2
r2_pacmap <- diff(lim2)/diff(lim1)

pacmap_two_non_linear_diff_shaped_close_clusters_plt <- pacmap_two_curvy_scaled |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("c1", sc_ltr_pos) +
  theme(aspect.ratio = 1)

## hexagon binning to have regular hexagons
hb_obj_pacmap_two_curvy2 <- hex_binning(
  data = pacmap_two_curvy_scaled, 
  bin1 = 8, 
  r2 = r2_pacmap)

a1_2_pacmap <- calc_bins_y(
  bin1 = 8, 
  r2 = r2_pacmap
)$a1

## Data set with all centroids
all_centroids_df4 <- hb_obj_pacmap_two_curvy2$centroids

## Generate all coordinates of hexagons
hex_grid4 <- hb_obj_pacmap_two_curvy2$hex_poly

## To obtain the standardise counts within hexbins
counts_df4 <- hb_obj_pacmap_two_curvy2$std_cts
df_bin_centroids4 <- extract_hexbin_centroids(
  centroids_df = all_centroids_df4, 
  counts_df = counts_df4) |>
  filter(drop_empty == FALSE) |>
  mutate(b1 = "b1 = 11")

pacmap_data_two_curvy2_with_hb_id <- hb_obj_pacmap_two_curvy2$data_hb_id
df_all_two_curvy4 <- dplyr::bind_cols(training_data_two_curvy, pacmap_data_two_curvy2_with_hb_id)
df_bin_two_curvy4 <- avg_highd_data(data = df_all_two_curvy4)

## Triangulate bin centroids
tr1_object_two_curvy4 <- tri_bin_centroids(
  df_bin_centroids4, x = "c_x", y = "c_y")
tr_from_to_df_two_curvy4 <- gen_edges(
  tri_object = tr1_object_two_curvy4)

trimesh_two_curvy_removed1_pacmap_with_data <- ggplot() +
  geom_segment(data = tr_from_to_df_two_curvy4,
               aes(
                 x = x_from,
                 y = y_from,
                 xend = x_to,
                 yend = y_to),
               colour = "#000000") +
  geom_point(data = pacmap_two_curvy_scaled,
             aes(
               x = PaCMAP1,
               y = PaCMAP2
             ),
             color = "#e41a1c",
             alpha = 0.5,
             size = 0.5) +
  coord_equal() +
  interior_annotation("c2", sc_ltr_pos)

```{r}
#| label: best-fit-pacmap-two-curvy-proj

names(model_df)[7:NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(6 + NCOL(projected_model_df)):NCOL(model_df)] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

two_curvy_proj_pacmap_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") + #31a354
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#e41a1c") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("c3", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )

two_curvy_proj_pacmap_all_model <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = true_model_df_proj, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#e41a1c") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.126, 0.125)) + 
  ylim(c(-0.126, 0.125)) +
  interior_annotation("c4", sc_ltr_pos) +
  theme(
    legend.position = "none"
  )
```

@@

nldr_two_curvy_original2 + tsne_two_non_linear_diff_shaped_close_clusters_plt +
  pacmap_two_non_linear_diff_shaped_close_clusters_plt +
  trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
trimesh_two_curvy_removed1_pacmap_with_data +
  two_curvy_proj_umap_model2 +
  two_curvy_proj_tsne_model +
  two_curvy_proj_pacmap_model +
  two_curvy_proj_umap_all_model +
  two_curvy_proj_tsne_all_model +
  two_curvy_proj_pacmap_all_model +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)")))
```

@@

@@

@@

tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds")

@@

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
pbmc_scaled_obj <- gen_scaled_data(
  data = umap_pbmc)
umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
```

>>>>

```{r}
#| label: code-illustration
# Code to draw illustration for notation
## hexagon binning to have regular hexagons
hb_obj_notation <- hex_binning(
  nldr_obj = algo_obj_two_curvy$nldr_obj, 
  bin1 = 7, 
  q = 0.1)

@@

hex_grid_coloured_two_curvy2 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy2, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", 
        linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = emb1, y = emb2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "C") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos, cex = 1.5)

@@

algo_obj_two_curvy3 <- fit_highd_model(
    highd_data = training_data_two_curvy, 
    nldr_data = tsne_two_curvy, 
    bin1 = num_bins_x_two_curvy, 
    q = 0.1, 
    benchmark_highdens = 5)

tr_from_to_df_two_curvy3 <- algo_obj_two_curvy3$trimesh_data
df_bin_centroids_two_curvy3 <- algo_obj_two_curvy3$model_2d
df_bin_two_curvy3 <- algo_obj_two_curvy3$model_highd

hex_grid_two_curvy3 <- algo_obj_two_curvy3$hb_obj$hex_poly
counts_df_two_curvy3 <- algo_obj_two_curvy3$hb_obj$std_cts

@@

hex_grid_coloured_two_curvy3 <-  ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts_two_curvy3, 
    aes(x = x, y = y, 
        group = hex_poly_id, 
        fill = std_counts), 
        color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
           aes(x = emb1, y = emb2),
           alpha = 0.3,
           size = 0.5) +
  scale_fill_viridis_c(direction = -1, na.value = "#ffffff", option = "D") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos, cex = 1.5)
```

@@

predict_two_curvy_obj <- gen_scaled_data(
    nldr_data = predict_tsne_two_curvy)

@@

```{r}
#| label: read-two-curvy-clust-nldr
# Read a variety of different NLDR representations of two_non_linear_diff_shaped_close_clusters
# and plot them on same aspect ratio
tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_30.rds")

nldr_two_curvy1 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("a", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

tsne_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_tsne_perplexity_62.rds")

nldr_two_curvy6 <- tsne_two_curvy |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("f", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

umap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_umap_n-neigbors_15_min-dist_0.1.rds")

nldr_two_curvy2 <- umap_two_curvy |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#636363') +
  interior_annotation("b", c(0.08, 0.93)) +
  theme(aspect.ratio = 1)

phate_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_phate_knn_5.rds")

nldr_two_curvy3 <- phate_two_curvy |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("c") +
  theme(aspect.ratio = 1)

trimap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_trimap_n-inliers_12_n-outliers_4_n-random_3.rds")

nldr_two_curvy4 <- trimap_two_curvy |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#4daf4a') +
  interior_annotation("d") +
  theme(aspect.ratio = 1)

pacmap_two_curvy <- read_rds("data/two_non_linear_diff_shaped_close_clusters/two_non_linear_diff_shaped_close_clusters_pacmap_n-neighbors_10_init_random_MN-ratio_0.5_FP-ratio_2.rds")

nldr_two_curvy5 <- pacmap_two_curvy |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e") +
  theme(aspect.ratio = 1)
```

```{r}
#| label: combine-data-two_curvy

error_two_curvy_umap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_umap.rds")
error_two_curvy_tsne <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne.rds")
error_two_curvy_phate <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_phate.rds")
error_two_curvy_trimap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_trimap.rds")
error_two_curvy_pacmap <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_pacmap.rds")
error_two_curvy_tsne2 <- read_rds("data/two_non_linear_diff_shaped_close_clusters/error_two_non_linear_diff_shaped_close_clusters_tsne2.rds")

error_two_curvy_tsne2 <- error_two_curvy_tsne2 |>
  mutate(method = "tSNE2")

error_two_curvy <- bind_rows(error_two_curvy_umap, 
                         error_two_curvy_tsne,
                         error_two_curvy_phate,
                         error_two_curvy_trimap,
                         error_two_curvy_pacmap,
                         error_two_curvy_tsne2)

error_two_curvy <- error_two_curvy |>
  mutate(a1 = round(a1, 2)) |>
  filter(bin1 >= 5) |>
  group_by(method, a1) |>
  filter(MSE == min(MSE)) |>
  ungroup()
```

```{r}
#| label: error-comp-two_curvy

error_plot_two_curvy <-
  plot_mse(error_two_curvy) +
  scale_x_continuous(breaks =
    sort(unique(error_two_curvy$a1))[
      seq(1, length(
        unique(error_two_curvy$a1)), 
        by = 5)]) +
  scale_color_manual(
    values=c('#e41a1c','#ff7f00','#4daf4a', 
             "#a65628",'#636363', '#984ea3')) +
  scale_y_log10() + 
  theme(aspect.ratio = 1.5)
```

@@

error_plot_two_curvy + wrap_plots(nldr_two_curvy1, nldr_two_curvy2, nldr_two_curvy3, 
                                    nldr_two_curvy4, nldr_two_curvy5,
                                    nldr_two_curvy6, ncol = 2, 
                                    widths = c(50, 50))
```

@@

two_curvy_tsne_scaled_obj <- gen_scaled_data(
  nldr_data = tsne_two_curvy)

@@

tsne_two_curvy_plt <- tsne_two_curvy_scaled |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b1", sc_ltr_pos) +
  theme(aspect.ratio = 1)

## hexagon binning to have regular hexagons
hb_obj_tsne_two_curvy2 <- hex_binning(
  nldr_obj = algo_obj_two_curvy$nldr_obj, 
  bin1 = 24)

a1_2_tsne <- hb_obj_tsne_two_curvy2$a1

@@

## To obtain the standardise counts within hexbins
counts_df4 <- hb_obj_tsne_two_curvy2$std_cts
df_bin_centroids4 <- extract_hexbin_centroids(
  centroids_data = all_centroids_df4, 
  counts_data = counts_df4) |>
  mutate(b1 = "b1 = 11")

@@

tsne_data_two_curvy2_with_hb_id <- hb_obj_tsne_two_curvy2$data_hb_id
df_bin_two_curvy4 <- avg_highd_data(highd_data = training_data_two_curvy,
                                    scaled_nldr_hexid = tsne_data_two_curvy2_with_hb_id)

@@

## Triangulate bin centroids
tr1_object_two_curvy4 <- tri_bin_centroids(
  df_bin_centroids4)
tr_from_to_df_two_curvy4 <- gen_edges(
  tri_object = tr1_object_two_curvy4,
  a1 = a1_2_tsne)

@@

names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_from")

model_df <- dplyr::left_join(model_df, projected_model_df, by = c("to" = "ID"))
names(model_df)[(length(names(model_df))-1):length(names(model_df))] <- paste0(names(projected_model_df)[-NCOL(projected_model_df)], "_to")

@@

@@

nldr_two_curvy_original2 + tsne_two_curvy_plt +
  pacmap_two_curvy_plt +
  trimesh_two_curvy_removed1_with_data + trimesh_two_curvy_removed1_tsne_with_data +
trimesh_two_curvy_removed1_pacmap_with_data +
  two_curvy_proj_umap_model2 +
  two_curvy_proj_tsne_model +
  two_curvy_proj_pacmap_model +
  two_curvy_proj_umap_all_model +
  two_curvy_proj_tsne_all_model +
  two_curvy_proj_pacmap_all_model +
  plot_layout(ncol = 3) +
  plot_annotation(tag_levels = list(c("(a)", "(b)", "(c)")))
```

@@

algo_obj_gau1 <- fit_highd_model(
  highd_data = training_data_gau, 
  nldr_data = umap_data_gau, 
  bin1 = num_bins_x_gau1, 
  q = 0.1, 
  benchmark_highdens = 5)

umap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
df_bin_gau1 <- algo_obj_gau1$model_highd

@@

algo_obj_gau1 <- fit_highd_model(
  highd_data = training_data_gau, 
  nldr_data = pacmap_data_gau, 
  bin1 = num_bins_x_gau1, 
  q = 0.1, 
  benchmark_highdens = 5)

pacmap_gau_scaled <- algo_obj_gau1$nldr_obj$scaled_nldr
tr_from_to_df_gau1 <- algo_obj_gau1$trimesh_data
df_bin_centroids_gau1 <- algo_obj_gau1$model_2d
df_bin_gau1 <- algo_obj_gau1$model_highd

@@

tsne_one_c_shaped <- read_rds(file = "data/one_c_shaped_dens_structure/one_c_shaped_dens_structure_tsne_perplexity_52.rds") |>
  mutate(ID = row_number())

@@

umap_pbmc <- read_rds("data/pbmc3k/pbmc_umap_30_min_dist_0.3.rds")
pbmc_scaled_obj <- gen_scaled_data(
  nldr_data = umap_pbmc)
umap_pbmc_scaled <- pbmc_scaled_obj$scaled_nldr
```
